### 2020-03-11 paper/sections/abstract.tex
add the paper framework

\begin{abstract}
With the growth of computer vision based applications and services, an explosive amount of images have been uploaded to cloud servers that host such computer vision algorithms, usually in the form of deep learning models. JPEG has been used as the {\em de facto} compression and encapsulation method before one uploads the images, due to its wide adaptation. However, standard JPEG configuration does not always perform well for compressing images that are to be processed by a deep learning model, e.g., the standard quality level of JPEG leads to 50\% of size overhead (compared with the best quality level selection) on ImageNet under the same inference accuracy in popular computer vision models including InceptionNet, ResNet, etc. Knowing this, designing a better JPEG configuration for online computer vision services is still extremely challenging: 1) Cloud-based computer vision models are usually a black box to end-users; thus it is difficult to design JPEG configuration without knowing their model structures. 2) JPEG configuration has to change when different users use it since different input image's features and backend deep learning models. In this paper, we propose a reinforcement learning based JPEG configuration framework. In particular, we design an agent that adaptively chooses the compression quality level according to the input image's features and backend deep learning models. Then we train the agent in a reinforcement learning way to adapt it for different deep learning cloud services that act as the {\em interactive training environment} and feeding a reward with comprehensive consideration of accuracy and data size. In our real-world evaluation on Amazon Rekognition, Face++ and Baidu Vision, our approach can reduce the size of images by 1/2 -- 1/3 while the overall classification accuracy only decreases slightly. To make the solution adaptive to the changing input images, we propose an explore-exploit mechanism to adapt the agent to different ``scenery'' online. In our experiment, once capturing the scenery change, AdaCompress either reloads or retrains the RL agent based on the performance of the cached model to make a proper compression strategy and maintain the inference accuracy.
\end{abstract}