### 2020-03-11 paper/sections/design.tex

\section{Detailed Design}
\label{sec: design}

\begin{figure*}
    %\begin{tabular}{cc}
    \begin{minipage}{\linewidth}
        \centerline{\includegraphics[width=0.8\linewidth]{figures/conventional-framework.pdf}}
        \begin{center}
            {(a) Conventional solution: fixed user-defined compression quality level}
        \end{center}
%        \vspace{0.3cm}
    \end{minipage}
    \vfill
    \vspace{0.4cm}
    \begin{minipage}{\linewidth}
        \centerline{\includegraphics[width=0.8\linewidth]{figures/adaptive-framework.pdf}}
        \vspace{0.2cm}
        \begin{center}
            {(b) AdaCompress solution: input image and model aware compression}
        \end{center}
    \end{minipage}
    %\end{tabular}
    \vspace{0.1cm}
    \caption{Comparing to the conventional solution, our solution can update the compression strategy based on the backend model feedback}
    \label{fig: framework}
\end{figure*}

A brief framework of AdaCompress is shown in Figure \ref{fig: framework}. Briefly, it is an RL (reinforcement learning) based system to train an agent to choose the proper compression quality level $ c $ for one image to be compressed by JPEG. We discuss the formulation, agent design, reinforcement learning framework, feedback reward, and inference-estimation-querying-retraining mechanism separately in the following subsections. We provide experimental details of all the hyperparameters in Sec. \ref{sec: evaluation}. %% \\

\subsection{Problem Formulation}
\label{subsec: formulation}

% \colorbox{red}{notation table}

Without loss of generality, we denote the cloud deep learning service as $ \vec{y}_i = M(x_i) $ that provides a predicted result list $ \vec{y}_i $ for each input image $ x_i $, and it has a baseline output $ \vec{y}_{\rm ref} = M(x_{\rm ref}) $ for each reference input $ x \in X_{\rm ref} $. We use this $ \vec{y}_{\rm ref} $ as the ground truth label, and for each image $ x_c $ compressed at compression quality level $ c $, the output $ \vec{y}_c = M(x_c) $. Therefore, we have an accuracy metric $ \mathcal{A}_c $ by comparing $ \vec{y}_{\rm ref} $ and $ \vec{y}_c $. In general, we use the top-5 accuracy as the following $ \mathcal{A} $, the same as the classification metric of ILSVRC2012~\cite{ILSVRC12}.

\begin{align}
\mathcal{A} =& \frac{1}{k} \sum_{k}\max_jd(l_j, g_k) \\
& l_j \in \vec{y}_c, \quad j=1,...,5 \\
& g_k \in \vec{y}_{\rm ref}, \quad k=1, ..., {\rm length}(\vec{y}_{\rm ref}) \\
& d(x, y) = 1 \ \text{if} \ x=y  \ \text{else} \ 0 
\end{align}

Where $ j = 1,...,5 $ indicates the prediction labels at top-5 score, $ k = 1,...,{\rm length}(\vec{y}_{\rm ref}) $ means that if anyone of the top-5 predicted labels matches one of the predictions from $ \vec{y}_{\rm ref} $, it would be regarded as a correct prediction. In general, we cannot get the cloud deep learning model's in-layer details (e.g., softmax probabilities) for a cloud deep learning service, therefore we use a binary hard label $ d(x, y) \in \{0, 1\} $ to evaluate the accuracy. %% $ d(x, y) \in \{0, 1\} $

We also denote JPEG input images as $ f_{ic} = J(x_i, c) $ that for an input image $ x_i $ and a given compression quality level $ c $, and it outputs a compressed file $ f_{ic} $ at the size of $ s_{ic} $. Especially for a reference compression quality level $ c_{\rm ref} $, the compressed file size is $ s_{\rm ref} $. Besides, the input image from a specific location usually belongs to a particular contextual group. For example, in daytime scenery, the input images are expected to have a bright background and high contrast while nighttime images are usually gray-scaled thermal images. Therefore, the agent in one place and one scenery does not need to know all the contextual features in all places and all sceneries. We formulate this as contextual group $ \mathcal{X} $. This contextual grouping concept is also discussed in the study ~\cite{mcdnn}.

Initially, the agent tries different compression quality levels $ c_{\min} < c < c_{\max}, c \in \mathbb{N} $ to obtain compressed image $ x_c $ from input image $ x $, and an image compressed at a reference level $ c_{\rm ref} $ is also uploaded to the cloud to obtain $ \vec{y}_{\rm ref} $. Comparing the two uploaded instances $ \{x_{\rm ref}, x_c\} $ and cloud recognition results $ \{\vec{y}_{\rm ref}, \vec{y}_c\} $, we can obtain the reference file size $ s_{\rm ref} $ and compressed file size $ s_c $, therefore compute the file compression ratio $ \Delta s = \frac{s_c}{s_{\rm ref}} $ and accuracy metric $ \mathcal{A}_c $.

\subsection{RL Agent Design}

The RL agent is expected to give a proper compression quality level $ c $ that minimizing the file size $ s_c $ while keeping the accuracy $ \mathcal{A} $. For the RL agent, the input features are continuous numerical vectors, and the expected output is discrete compression quality level $ c $, therefore we can use Deep Q-learning Network \cite{DQN} as the RL agent. But naive Deep Q-learning Network can't work well in this task because of the following challenges: %% \\

\begin{itemize}
    \item The state space of reinforcement learning is too large, and to preserve enough details, we have to add many layers and nodes to the neural network, making the RL agent extremely difficult to converge. 
    \item It takes a long time to train one step in a large inference neural network, making the training process too time-consuming.
    \item The RL agent starts training from random trials and starts learning after it finds better feedback reward. When training from a randomly initialized neural network, the feedback reward is very sparse, making it difficult for the agent to learn.
\end{itemize}

To address these challenges, we use the early layers of a well-trained neural network to extract the structural information of an input image. This is a commonly used strategy in training a deep neural network~\cite{finetunning,finetunning2}. Therefore instead of training an RL agent directly from the input image, we use a pre-trained small neural network to extract the features from the input image to reduce the input dimension and accelerate the training procedure. In this work, we use the early convolution layers of MobileNetV2~\cite{MobileNetV2} as the image feature extractor $ \mathcal{E}(\cdot) $ for its efficiency in image classification. The Deep Q-learning Network $ \phi $ is connected to the feature extractor's last convolution layer. We update the RL agent's policy by changing the parameters of the Deep Q-learning Network $ \phi $ while the feature extractor $ \mathcal{E} $ remains fixed. %% \\

\subsection{Reinforcement Learning Framework}

In a specific scenery where the user input image $ x $ belongs context group $ \mathcal{X} $, we define the contextual information $ \mathcal{X} $, along with the backend cloud model $ M $, as the \emph{emulator environment} $ \{\mathcal{X}, M\} $ of the reinforcement learning problem.

Based on this insight, we formulate the feature extractor's output $ \mathcal{E}(J(\mathcal{X}, c)) $ as \emph{states} and the compression quality level $ c $ as discrete \emph{actions}. In our system, to accelerate training, we define 10 discrete actions to indicate 10 compression quality levels of JPEG ranging from $ 5, 15, ...,95 $. We denote the \emph{action-value function} as $ Q(\phi(\mathcal{E}(f_t)), c; \theta) $, then the optimal compression quality level at time $ t $ is $ c_t = {\rm argmax}_cQ(\phi(\mathcal{E}(f_t)), c; \theta) $ where $ \theta $ indicates the parameters of the Deep Q-learning Network $ \phi $. In such reinforcement learning formulation, the training phase is to minimize a loss function $ L_i(\theta_i) = \mathbb{E}_{s, c \sim \rho (\cdot)}\Big[\big(y_i - Q(s, c; \theta_i)\big)^2 \Big] $ that changes at each iteration $ i $ where $ s = \mathcal{E}(f_t) $, and $ y_i = \mathbb{E}_{s' \sim \{\mathcal{X}, M\}} \big[ r + \gamma \max_{c'} Q(s', c'; \theta_{i-1}) \mid s, c \big] $ is the target for the iteration $ i $. Especially, $ r $ is the feedback reward, and $ \rho(s, c) $ is a probability distribution over sequences $ s $ and compression quality level $ c $~\cite{DQN}. When minimizing the distance of the action-value function's output $ Q(\cdot) $ and target $ y_i $, the action-value function $ Q(\cdot) $ outputs a more accurate estimation of an action.

In such a formulation, it is similar to the Deep Q-learning Network problem but not the same. Different from conventional reinforcement learning, the interactions between the agent and environment are infinite, therefore there is no signal from the environment telling that an episode has finished. Therefore, we train the RL agent intermittently at a manual interval of $ T $ after the condition $ t \geq T_{\rm start} $ guaranteeing that there are enough transitions in the memory buffer $ \mathcal{D} $. In the training phase, the RL agent firstly takes some random trials to observe the environment's reaction, and we decrease the randomness when training. All transitions are saved into a memory buffer $ \mathcal{D} $, and the agent learns to optimize its action by minimizing the loss function $ L $ on a mini-batch from $ \mathcal{D} $. The training procedure would converge as the agent's randomness keeps decaying. Finally, the agent's action is based on its historical optimal experiences. The training procedure is presented in Algorithm \ref{alg: rl-train}, and we list the parameters in Sec.~\ref{sec: evaluation}.

\begin{algorithm}[htbp]
    \caption{Training RL agent $ \phi $ in environment $ \{\mathcal{X}, M\} $}
    \label{alg: rl-train}
    \begin{algorithmic}[1]
        \STATE Initialize replay memory buffer $ \mathcal{D} $ to capacity $ N $
        \STATE Initialize action-value function $ Q $ with random weights $ \theta $
        \STATE Initialize sequence $ s_1 = \mathcal{E}\big(J(x_1, c_1)\big), x_1 \in \mathcal{X} $ and $ \phi_1 = \phi(f_1) $
        \FOR {t = 1, $ K $}
        \STATE With probability $ \epsilon $ select a random compression level $ c_t $ otherwise select $ c_t = {\rm argmax}_cQ\Big(\phi\big(\mathcal{E}(f_t)\big), c; \theta\Big) $
        \STATE Compress image $ x_t $ at quality $ c_t $ and upload it to the cloud to get result $ (\vec{y}_{\rm ref}, \vec{y}_c) $ and calculate reward $ r = R(\Delta s, \mathcal{A}_c) $
        \STATE Set $ s_{t+1} = s_t $, generate $ c_t, x_{t+1} $ and preprocess $ \phi_{t+1} = \phi \big(\mathcal{E}(f_{t+1})\big) $
        \STATE Store transition $ (\phi_t, c_t, r_t, \phi_{t+1}) $ in $ \mathcal{D} $
        \IF {$ t \mod T == 0 $ and $ t \geq T_{\rm start} $}
        \STATE Sample randomly mini-batch of transitions \\ $ (\phi_j, c_j, r_j, \phi_{j+1}) $ from memory buffer $ \mathcal{D} $
        \STATE Set $ y_i = r_j + \gamma \max_{c'}Q(\phi_{j+1}, c'; \theta) $
        \STATE Compute decay exploration rate \\
        $ \epsilon = 
        \begin{cases}
        \mu_{\rm dec}\cdot \epsilon & \text{ if } \ \mu_{\rm dec}\cdot \epsilon > \epsilon_{\min} \\ 
        \epsilon_{\min}             & \text{ if } \ \mu_{\rm dec}\cdot \epsilon \leq \epsilon_{\min}
        \end{cases} $
        \STATE Perform a gradient descent step on \\ $ \big(y_j - Q(\phi_j, c_j; \theta)\big)^2 $ according to ~\cite{DQN}
        \ENDIF
        \ENDFOR
    \end{algorithmic}
\end{algorithm}

\subsection{Feedback Reward Design}

In our solution, the agent is trained by the feedback reward from the environment $ \{\mathcal{X}, M\} $. In the above formulation, we defined compression rate $ \Delta s = \frac{s_c}{s_{\rm ref}} $ and accuracy metric $ \mathcal{A}_c $ in compression quality level $ c $. Basically, we want the agent to choose a proper compression quality level that minimizing the file size while remaining acceptable accuracy, therefore the overall reward $ r $ should be in proportion to the accuracy $ \mathcal{A} $ while in inverse proportion to the compression ratio $ \Delta s $. We introduce two linear factors $ \alpha $ and $ \beta $ to form a linear combination $ r = \alpha \mathcal{A} - \Delta s + \beta $ as the reward function $ R(\Delta s, \mathcal{A}) $. %% \\

\subsection{Inference-Estimation-Querying-Retraining Mechanism}

As a running system, we introduce an inference-estimation-querying-retraining mechanism to cope with the scenery change in the inference phase, building a system with different components to inferring, capturing the scenery change, then either reloading or retraining the RL agent based on the performance of the cached model. The overall system diagram is illustrated in Figure \ref{fig: diagram}.

%\begin{figure}[H]
%    \includegraphics[width=\linewidth]{figures/overall-diagram.pdf}
%    \caption{Diagram of AdaCompress architecture}
%    \label{fig: diagram}
%\end{figure}

\begin{figure*}[htbp]
    \centerline{\includegraphics[width=0.8\linewidth]{figures/overall-diagram.pdf}}
    \vspace{0.2cm}
    \caption{Diagram of AdaCompress architecture}
    \label{fig: diagram}
\end{figure*}

%% The system diagram is shown in Figure \ref{fig: diagram}.
We build up the memory buffer $ \mathcal{D} $ and RL (reinforcement learning) training kernel based on the compression and upload driver. When the RL kernel is called, it will load transitions from the memory buffer $ \mathcal{D} $ to train the compression quality level predictor $ \phi $. When the system is deployed, the pre-trained RL agent $ \phi $ guides the compression driver to compress the input image with an adaptive compression quality level $ c $ according to the input image, then uploads the compressed image to cloud. %% \\

After the AdaCompress is deployed, the input images scenery context $ \mathcal{X} $ may change (e.g., day to night, sunny to rainy), when the scenery changes, the older RL agent's compression selection strategy may not be suitable anymore, causing the overall accuracy decreases. To cope with the scenery change, we invoke an estimator with a probability $ p_{\rm est} $. We do this by generating a random value $ \xi \in (0,1) $ and comparing it to $ p_{\rm est} $. If $ \xi \leq p_{\rm est} $ the estimator would be invoked, and AdaCompress would upload the reference image $ x_{\rm ref} $ along with the compressed image $ x_i $ to obtain $ \vec{y}_{\rm ref} $ and $ \vec{y}_i $, therefore calculate $ \mathcal{A}_i $, and save the transition $ (\phi_i, c_i, r_i, \mathcal{A}_i) $ to the memory buffer $ \mathcal{D} $. The estimator also compares the recent $ n $ steps' average accuracy $ \bar{\mathcal{A}}_n $ and the initial accuracy threshold $ \mathcal{A}_0 $ in memory $ \mathcal{D} $. Once the recent average accuracy is much lower than the initial accuracy threshold, the estimator would query an RL agent model to replace the current agent model and test the performance of the reloaded model. To test the reloaded model's performance, the estimator computes the average accuracy $ \bar{\mathcal{A}}^*_n $. If $ \bar{\mathcal{A}}^*_n $ is still lower than the accuracy threshold, the estimator would invoke the RL training kernel to retrain the agent. Once the estimator discovers that the trained reward is higher than the reward threshold, it would stop the training kernel, cache the trained RL agent and switch back to the normal inference state.

Since the reference image $ x_{\rm ref} $ and the compressed image $ x_i $ are both needed in the retraining phase, causing a large uploading file size overhead, especially when the scenery changes frequently (e.g., day to night, then night to day). To avoid unnecessary upload traffic load in the retraining phase, we build up the model caching library to cache the trained RL agent models. When capturing the scenery change, we query a pre-trained model from the model caching library rather than retraining from scratch. %% \\

AdaCompress adaptively switch itself between four states. The switching policy is shown as Figure~\ref{fig: state-switching}.

%\begin{figure}[H]
%    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.8cm, semithick]
%    \tikzstyle{every state}=[ellipse, align=center, draw=blue, text=black]
%    
%    \node[initial, state] (B)                    {inference};
%    \node[state]         (C) [below right of=B] {retrain};
%    \node[state]         (D) [above right of=C] {estimate};
%    
%    \path   (B) edge [loop above] node {$ \xi > p_{\rm est} $} (B)
%    (B.10)    edge                node {$ \xi \leq p_{\rm est} $} (D.170)
%    (C) edge [below left]      node {$ \bar{r}_n > r_{\rm th} $} (B)
%    edge [loop below] node {$ \bar{r}_n \leq r_{\rm th} $}     (C)
%    (D) edge [loop right] node {$ \xi \leq p_{\rm est} $} (D)
%    (D.190)    edge              node {$ \xi > p_{\rm est} $} (B.-10)
%    (D.225)    edge [below right]      node {$ \bar{\mathcal{A}_n} < \mathcal{A}_0 $} (C);
%    \end{tikzpicture}
%    % \vspace{-0.3cm}
%    \caption{State switching policy}
%    \label{fig: state-switching}
%    %\vspace{-0.8cm}
%\end{figure}

\begin{figure}[H]
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.1cm, semithick]
    \tikzstyle{every state}=[ellipse, align=center, draw=blue, text=black]

    \node[initial, state] (B)                    {inference};
    \node[state]         (C) [below of=B] {retraining};
    \node[state]         (D) [right of=B, node distance=4cm] {estimation};
    \node[state]         (E) [right of=C, node distance=4cm] {model querying};
%    \node[state][rectangle]   (F) [below right of=C, node distance=2.5cm] {Model Library};

    \path   (B) edge [loop above] node {$ \xi > p_{\rm est} $} (B)
    (B.10)    edge                node {$ \xi \leq p_{\rm est} $} (D.170)
    (C) edge [below left]      node {$ \bar{r}_n > r_{\rm th} $} (B)
    edge [loop below] node {$ \bar{r}_n \leq r_{\rm th} $}     (C)
    (D) edge [loop above] node {$ \xi \leq p_{\rm est} $} (D)
    (D.190)    edge              node {$ \xi > p_{\rm est} $} (B.-10)
    (D)    edge [below right]      node {$ \bar{\mathcal{A}_n} < \mathcal{A}_0 $} (E)
    (E) edge [below left] node [left] {$ \bar{\mathcal{A}^*_n} \geq \mathcal{A}_0 $} (B)
    (E) edge [below left] node [below] {$ \bar{\mathcal{A}^*_n} < \mathcal{A}_0 $} (C);
%    (C) edge [below right] node [left] {cache model} (F)
%    (F) edge [below right] node [right] {load model} (E);
    \end{tikzpicture}
    % \vspace{-0.3cm}
    \caption{State switching policy}
    \label{fig: state-switching}
    %\vspace{-0.8cm}
\end{figure}

\subsubsection{\textbf{Inference:}}

For most times, AdaCompress runs in this state. In this state, only the compressed images are uploaded to the cloud to achieve minimum uploading file size overhead. To keep a stable accuracy performance even the input image scenery changes, the agent would occasionally switch to the estimation state with the probability of $ p_{\rm est} $, meanwhile remain the inference state with the probability of $ 1 - p_{\rm est} $. %% \\

\subsubsection{\textbf{Estimation:}}

In this state, the reference image $ x_{\rm ref} $ and compressed image $ x_i $ are uploaded to the cloud simultaneously to obtain $ \vec{y}_{\rm ref} $ and $ \vec{y}_i $ which are used to calculate $ \mathcal{A}_i $. In each epoch $ i $ the transition $ (\phi_i, c_i, r_i, \mathcal{A}_i) $ is logged in a memory buffer $ \mathcal{D} $. When the average accuracy $ \bar{\mathcal{A}}_n $ of the latest $ n $ steps is higher than the accuracy threshold $ \mathcal{A}_0 $, it stays in the estimation state with the probability of $ p_{\rm est} $ or switches back into the inference state with the probability of $ 1 - p_{\rm est} $. Once the average accuracy $ \bar{\mathcal{A}}_n $ is lower than the initial accuracy threshold $ \mathcal{A}_0 $, indicating that the current agent is no more suitable for the current input image scenery, AdaCompress turns into the model querying state and loads a new RL agent model from the model caching library. %% \\

Therefore, the estimating probability of $ p_{\rm est} $ is vital to the whole system. On the one hand, the estimator should be invoked occasionally to estimate current agent's accuracy, so that to capture the scenery change on time; on the other hand, the estimator uploads the reference image $ x_{\rm ref} $ along with the compressed image, therefore the uploading file size overhead is greater than the conventional benchmark solution, causing higher upload traffic load. %% \\

To achieve trade-off between the risk of the scenery change and the objective of reducing upload traffic load, we design an accuracy-aware dynamic $ p_{\rm est} $ solution. We first define that after running for $ N $ steps, the recent $ n $ steps' average accuracy is:

\begin{align*}
\bar{\mathcal{A}_n} &=
\begin{cases}
\frac{1}{n}\sum_{i=N-n}^{N} \mathcal{A}_i & \text{ if } N \geq n \\ 
\frac{1}{n}\sum_{i=1}^{n} \mathcal{A}_i & \text{ if } N < n 
\end{cases}
\end{align*}

With this definition, an intuitive formulation of the change of $ p_{\rm est} $ is in inverse proportion of the gradient of $ \mathcal{A} $, meaning that when the recent accuracy is going down, the estimation probability $ p_{\rm est} $ would increase. We formulate that $ p'_{\rm est} = p_{\rm est} + \omega \nabla \bar{\mathcal{A}} $ where $ \omega $ is a scaling factor. With this recursive formula, we have the general term of $ p_{\rm est} $ with an initial estimation probability $ p_0 $ is $p_{\rm est} = p_0 + \omega \sum_{i=0}^{N} \nabla \bar{\mathcal{A}_i} $.

\subsubsection{\textbf{Model Querying:}}

The model querying state is designed to cope with the scenery change before retraining. It loads a new RL agent model from the model caching library and tests whether the loaded agent model is suitable for the current scenery by re-calculating the average accuracy. If the new average accuracy $ \bar{\mathcal{A}}^*_n $ is higher than the accuracy threshold $ \mathcal{A}_0 $, indicating that the loaded agent model is suitable for the current scenery, AdaCompress would switch back to the normal inference state using the loaded agent model. Otherwise, AdaCompress would switch into the retraining state and invokes the RL training kernel to retrain a new RL agent for the current scenery.

In this way, AdaCompress is capable to cope with the scenery change in a cached ``memory'' manner, avoiding retraining the agent model at every scenery change, therefore cutting down the uploading file size overhead. Meanwhile, the agent caching strategy would cause a little storage overhead on local terminal devices.

\subsubsection{\textbf{Retraining:}}

This state is to adapt the agent to the current input image scenery by retraining it with the memory buffer $ \mathcal{D} $, which is similar to the training procedure. The retraining phase has finished upon the recent $ n $ steps' average reward $ \bar{r}_n $ is higher than the user-defined threshold $ r_{\rm th} $. And when the retraining procedure finishes, the memory buffer $ \mathcal{D} $ would be flushed, preparing to save new transitions for the retraining of a next scenery change. The trained RL agent model would be cached in the model caching library and be used to switch to the inference state.

\subsection{Insight of RL Agent's Behavior}
\label{subsec:insight}

In the inference phase, the pre-trained RL agent predicts a proper compression quality level according to the input image's feature. The reference image is not uploaded to the cloud anymore; only the compressed image is uploaded, therefore, the upload traffic load is reduced. We notice that the RL agent's behaviors are various for different input image sceneries and backend cloud services, therefore we try to make further investigations by plotting the RL agent's "attention map" (i.e., visual explanations of why the agent chooses a compression quality level). %% \\

\subsubsection{\textbf{Compression Level Choice Variation:}}

In our experiment, we find that in different cloud application environments, the agent's final chosen compression quality levels can be quite different. As shown in Figure \ref{fig: quality_chosen}, for Face++ and Amazon Rekognition, the agent's choices are concentrated at around $ c=15 $, but for Baidu Vision, the agent's choices are distributed more evenly. Therefore, the optimal compression strategy should be different for different backend cloud services. This variation is caused by the interaction between the agent and the backend cloud model in the training phase. Since the agent's training procedure is based on a specific backend cloud model $ M_1 $, for another backend cloud model $ M_2 $, the interaction between the agent and $ M_2 $ is quite different. Therefore the agent's best compression quality level selection presents variation for different backend cloud models.

\begin{figure}[htb]
    \includegraphics[width=\linewidth]{figures/quality_chosen.pdf}
    \caption{Histogram of RL agent's best compression quality level selection for different cloud services}
    \label{fig: quality_chosen}
    % \vspace{-0.3cm}
\end{figure}

Moreover, in our experiment, the agent presents different behaviors when the input images change from one dataset to another. Figure \ref{fig: dataset_change} shows the agent's choices for the same backend cloud model (Baidu Vision) but different image datasets. We prepare two datasets indicating two contextual sceneries. We randomly sample images from ImageNet~\cite{ImageNet} whose images are mostly taken in the daytime, to act as a daytime scenery, and randomly select nighttime images from the FLIR Thermal Dataset~\cite{FLIR} to form another dataset to act as a nighttime scenery. The histogram shown in Figure~\ref{fig: dataset_change} points out that, for the ImageNet images, the agent prefers a lower compression quality level, but its choices are distributed more evenly. For the FLIR Thermal images, the agent's choices are more accumulated in some relatively high compression quality levels. We can see that, to maintain high accuracy, when the input image's contextual group $ \mathcal{X} $ changes, the agent's compression quality level selection changes as well. This phenomenon presents that the agent can adaptively choose a proper compression quality level based on the input image's features. %% \\

%\begin{figure}[H]
%    \includegraphics[width=\linewidth]{figures/dataset_change.pdf}
%    \caption{Histogram of RL agent's best compression level selection for different scenery image inputs}
%    \label{fig: dataset_change}
%\end{figure}

\begin{figure}[htb]
    \includegraphics[width=\linewidth]{figures/dataset_change.eps}
    \caption{Histogram of RL agent's best compression quality level selection for different input image sceneries}
    \label{fig: dataset_change}
\end{figure}

\subsubsection{\textbf{Attention Map Variation:}}
\label{subsec: attention map}

To take insight investigations, we plot the importance map of a chosen compression quality level. We leverage a conventional visualization algorithm, Grad-Cam, to observe the Deep Q-Learning Network's interest when choosing compression quality levels. Grad-Cam is a widely used solution to present the importance map of a deep neural network, it is done by calculating the gradients of each target concept and backtracking to the final convolution layer. In this work, we plot the RL agent's attention map by Grad-Cam in Figure~\ref{fig: attention}. %% \\

\begin{figure*}[htbp]
    \begin{minipage}{0.2\linewidth}
        \centerline{\includegraphics[width=4.0cm, height=3.0cm]{figures/robust_figure1.png}}
        \centerline{(1a) Q=5}
    \end{minipage}
    \hfill
    \begin{minipage}{0.2\linewidth}
        \centerline{\includegraphics[width=4.0cm, height=3.0cm]{figures/robust_figure2.png}}
        \centerline{(1b) Q=15}
    \end{minipage}
    \hfill
    \begin{minipage}{0.2\linewidth}
        \centerline{\includegraphics[width=4.0cm, height=3.0cm]{figures/robust_figure3.png}}
        \centerline{(1c) Q=15}
    \end{minipage}
    \hfill
    \begin{minipage}{0.2\linewidth}
        \centerline{\includegraphics[width=4.0cm, height=3.0cm]{figures/robust_figure4.png}}
        \centerline{(1d) Q=15}
    \end{minipage}

    \vfill
    \vspace{0.4cm}

    \begin{minipage}{0.2\linewidth}
        \centerline{\includegraphics[width=4.0cm, height=3.0cm]{figures/sensetive_figure1.png}}
        \centerline{(2a) Q=85}
    \end{minipage}
    \hfill
    \begin{minipage}{0.2\linewidth}
        \centerline{\includegraphics[width=4.0cm, height=3.0cm]{figures/sensetive_figure2.png}}
        \centerline{(2b) Q=85}
    \end{minipage}
    \hfill
    \begin{minipage}{0.2\linewidth}
        \centerline{\includegraphics[width=4.0cm, height=3.0cm]{figures/sensetive_figure3.png}}
        \centerline{(2c) Q=75}
    \end{minipage}
    \hfill
    \begin{minipage}{0.2\linewidth}
        \centerline{\includegraphics[width=4.0cm, height=3.0cm]{figures/sensetive_figure4.png}}
        \centerline{(2d) Q=75}
    \end{minipage}
    \vspace{0.2cm}
    \caption{Visualization of the importance map for the RL agent to choose a compression quality level}
    \label{fig: attention}
\end{figure*}

In our investigations, we find that in different environments $ \{\mathcal{X}, M\} $, the RL agent picks up compression quality level based on the visual textures of different regions in the image. As shown in Figure \ref{fig: attention}, picture 1a -- 1d are some pictures that the agent chooses to compress highly, the agent selects lower compression quality levels based on the complex texture of the images. On the contrary, for pictures 2a -- 2d, the agent chooses higher compression quality levels to preserve more details, and the agent's interest falls on some smooth regions. Especially for 1a and 2a, in picture 1a, the agent chooses a lower quality compression level based on the rough central region though there are smooth regions around it, and in picture 2a, the agent chooses a relatively higher compression quality level based on the surrounding smooth region rather than the central region. %% \\